### 2018-05-15 submission/manuscript.Rmd
remove sub-sampling and fix cycle jargon

***The Number of OTUs are Dependent on HiFi DNA Polymerase:***
A consistent difference in the range of number of OTUs, that was dependent on the HiFi DNA polymerase used was observed for fecal samples [Figure 1]. Additionally, there was a trend for lower cycle numbers (15-20) to result in a reduced range in the number of OTUs versus higher cycle numbers (25, 30, and 35) between HiFi DNA polymerases [Figure 1]. For fecal samples, there was a significant difference between HiFi DNA polymerases at 35 cycles (F-stat > `r round(min(filter(fs_anova_summary, cycle == "35x", sub_sample_level == 1000)[, "F.value"]), digits = 2)`, P-value = `r format(filter(fs_anova_summary, cycle == "35x", sub_sample_level == 1000)[, "bh"], digits = 2, nsmall = 2, scientific = T)`) [Table S1]. No differences between the specific groups was found based on a Tukey post-hoc test (P-value `r ifelse(min(filter(fs_tukey_summary, sub_sample_level == 1000)[, "padj"]) > 0.05,  invisible(">"), invisible("check data"))` 0.05) [Table S2]. This HiFi DNA polymerase dependent difference in the range of number of OTUs also was observed in the mock community samples [Figure 2]. Regardless if fecal or mock communities were used, the same HiFi DNA polymerases had high (Platinum) and low (Accuprime) number of OTUs and this was consistent across cycle number and sub-sampling depth [Figure 1-2 & Table S1-S4]. In contrast to the results obtained with fecal samples, differences between HiFi DNA polymerases were observed as early as 20 cycles in the mock community samples (F-stat = `r round(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "F.value"], digits = 2)`, P-value = `r format(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "bh"], digits = 1, nsmall = 1)`) [Table S3]. For the mock community, the majority of  differences in the number of OTUs were between Kappa and Platinum and the other HiFi DNA polymerases across different number of cycles [Table S4]. Based on these observations in fecal and mock communities, it is clear that different HiFi DNA polymerases result in a different total number of OTUs within a sample.

***Minimal Bray-Curtis Differences are Detected and are Dependent on Cycle Number:***
A few small differences based on sub-sampling and cycle number were detected in overall bacterial community composition. Within the same fecal sample and independent of HiFi DNA polymerases, there were large differences in the community composition between 20 and 25 cycles [Figure 3A]. Additionally, when data was available for the mock communities, there were also large observed differences between 20 and 25 cycles [Figure 3B]. These community differences within a sample disappear when comparing 25 to 30 cycles and do not persist past 25 cycles [Figure 3]. Although these trends are clearly noticeable, we found that there was no detectable difference in Bray-Curtis index when comparing to the previous 5-cycle increment for both fecal and mock communities after multiple comparison correction (P-value `r ifelse(min(filter(bray_sim_to_prev_cycle_kruskal_results, sub_sample_level == 1000)[, "bh"], na.rm = T) > 0.05, invisible(">"), invisible("check data"))` 0.05). Using PERMANOVA to test for community differences based on cycle number within HiFi DNA polymerases, only Phusion had cycle dependent differences (P-value = `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 1000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)`. For fecal samples, Phusion was one of two HiFi DNA polymerases that had enough sequences to be rarefied to 1000 at 15 cycles. Overall, these data suggest that there are small HiFi DNA polymerase differences in Bray-Curtis index that are dependent on cycle number.

***Sequence Error is Dependent on both HiFi DNA Polymerase and Cycle Number:***
Differences in the median average per base error varied by HiFi DNA polymerase without a clear pattern across sub-sampling depth [Table S5]. The highest median average per base error rates were for the Kappa HiFi DNA polymerase [Figure 4]. This error rate was minimally affected by both the 'pre.cluster' step and chimera removal by VSEARCH [Figure 4]. The differences in the median average per base error rate between the different HiFi DNA polymerases was cycle dependent with Platinum having the largest changes versus other HiFi DNA polymerases [Figure 4B-C and Table S6]. The total sequences with at least one error was also cycle number dependent and differences between HiFi DNA polymerases could be drastically reduced by the use of the 'pre.cluster' step [Figure S1]. These differences in sequences with at least one error were mostly due to differences in Accuprime^TM^ and Platinum versus the other HiFi DNA polymerases [Figure S1 & Table S7 & S8]. Finally, we did not observe a HiFi DNA polymerase dependent difference on base substitution rate [Figure S2]. Although sequence error is dependent on HiFi DNA polymerase some of these error dependent differences can be corrected using existing bioinformatic approaches.

>>>>

***The Number of OTUs are Dependent on HiFi DNA Polymerase:***
A consistent difference in the range of number of OTUs, that was dependent on the HiFi DNA polymerase used was observed for fecal samples [Figure 1]. Additionally, there was a trend for lower number of cycles (15-20) to result in a reduced range in the number of OTUs versus higher number of cycles (25, 30, and 35) between HiFi DNA polymerases [Figure 1]. For fecal samples, there was a significant difference between HiFi DNA polymerases at 35 cycles (F-stat > `r round(min(filter(fs_anova_summary, cycle == "35x", sub_sample_level == 1000)[, "F.value"]), digits = 2)`, P-value = `r format(filter(fs_anova_summary, cycle == "35x", sub_sample_level == 1000)[, "bh"], digits = 2, nsmall = 2, scientific = T)`) [Table S1]. No differences between the specific groups was found based on a Tukey post-hoc test (P-value `r ifelse(min(filter(fs_tukey_summary, sub_sample_level == 1000)[, "padj"]) > 0.05,  invisible(">"), invisible("check data"))` 0.05) [Table S2]. This HiFi DNA polymerase dependent difference in the range of number of OTUs also was observed in the mock community samples [Figure 2]. Regardless if fecal or mock communities were used, the same HiFi DNA polymerases had high (Platinum) and low (Accuprime) number of OTUs and this was consistent across the number of cycles used [Figure 1-2 & Table S1-S4]. In contrast to the results obtained with fecal samples, differences between HiFi DNA polymerases were observed as early as 20 cycles in the mock community samples (F-stat = `r round(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "F.value"], digits = 2)`, P-value = `r format(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "bh"], digits = 1, nsmall = 1)`) [Table S3]. For the mock community, the majority of  differences in the number of OTUs were between Kappa and Platinum and the other HiFi DNA polymerases across different number of cycles [Table S4]. Based on these observations in fecal and mock communities, it is clear that different HiFi DNA polymerases result in a different total number of OTUs within a sample.

***Minimal Bray-Curtis Differences are Detected and are Dependent on Cycle Number:***
A few small differences based on the number of cycles were detected in overall bacterial community composition. Within the same fecal sample and independent of HiFi DNA polymerases, there were large differences in the community composition between 20 and 25 cycles [Figure 3A]. Additionally, when data was available for the mock communities, there were also large observed differences between 20 and 25 cycles [Figure 3B]. These community differences within a sample disappear when comparing 25 to 30 cycles and do not persist past 25 cycles [Figure 3]. Although these trends are clearly noticeable, we found that there was no detectable difference in Bray-Curtis index when comparing to the previous 5-cycle increment for both fecal and mock communities after multiple comparison correction (P-value `r ifelse(min(filter(bray_sim_to_prev_cycle_kruskal_results, sub_sample_level == 1000)[, "bh"], na.rm = T) > 0.05, invisible(">"), invisible("check data"))` 0.05). Using PERMANOVA to test for community differences based on the number of cycles within HiFi DNA polymerases, only Phusion had cycle dependent differences (P-value = `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 1000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)`. For fecal samples, Phusion was one of two HiFi DNA polymerases that had enough sequences to be rarefied to 1000 at 15 cycles. Overall, these data suggest that there are small HiFi DNA polymerase differences in Bray-Curtis index that are dependent on the number of cycles.

***Sequence Error is Dependent on both HiFi DNA Polymerase and Cycle Number:***
Differences in the median error rate varied by HiFi DNA polymerase across the number of cycles used [Table S5]. The highest error rates were for the Kappa HiFi DNA polymerase [Figure 4]. The differences in the median error rate between the different HiFi DNA polymerases was dependent on the number of cycles with Platinum and Kappa having the most differences versus other HiFi DNA polymerases [Figure 4 and Table S6]. The total sequences with at least one error was also cycle number dependent [Table S7 & S8. These differences in sequences with at least one error were mostly due to differences in Accuprime^TM^ and Kappa versus the other HiFi DNA polymerases [Table S7 & S8]. Finally, we did not observe a HiFi DNA polymerase dependent difference on base substitution rate [Figure S1]. These results suggest that sequence error is dependent on HiFi DNA polymerase. We next wanted to assess whether chimeras also were dependent on HiFi DNA polymerase and whether this affected OTU number.

### 2018-05-15 submission/manuscript.tex
remove sub-sampling and fix cycle jargon

\textbf{\emph{The Number of OTUs are Dependent on HiFi DNA Polymerase:}}
A consistent difference in the range of number of OTUs, that was
dependent on the HiFi DNA polymerase used was observed for fecal samples
{[}Figure 1{]}. Additionally, there was a trend for lower cycle numbers
(15-20) to result in a reduced range in the number of OTUs versus higher
cycle numbers (25, 30, and 35) between HiFi DNA polymerases {[}Figure
1{]}. For fecal samples, there was a significant difference between HiFi
DNA polymerases at 35 cycles (F-stat \textgreater{} 16.35, P-value =
9.7e-05) {[}Table S1{]}. No differences between the specific groups was
found based on a Tukey post-hoc test (P-value \textgreater{} 0.05)
{[}Table S2{]}. This HiFi DNA polymerase dependent difference in the
range of number of OTUs also was observed in the mock community samples
{[}Figure 2{]}. Regardless if fecal or mock communities were used, the
same HiFi DNA polymerases had high (Platinum) and low (Accuprime) number
of OTUs and this was consistent across cycle number and sub-sampling
depth {[}Figure 1-2 \& Table S1-S4{]}. In contrast to the results
obtained with fecal samples, differences between HiFi DNA polymerases
were observed as early as 20 cycles in the mock community samples
(F-stat = 15.82, P-value = 0.002) {[}Table S3{]}. For the mock
community, the majority of differences in the number of OTUs were
between Kappa and Platinum and the other HiFi DNA polymerases across
different number of cycles {[}Table S4{]}. Based on these observations
in fecal and mock communities, it is clear that different HiFi DNA
polymerases result in a different total number of OTUs within a sample.

\textbf{\emph{Minimal Bray-Curtis Differences are Detected and are
Dependent on Cycle Number:}} A few small differences based on
sub-sampling and cycle number were detected in overall bacterial
community composition. Within the same fecal sample and independent of
HiFi DNA polymerases, there were large differences in the community
composition between 20 and 25 cycles {[}Figure 3A{]}. Additionally, when
data was available for the mock communities, there were also large
observed differences between 20 and 25 cycles {[}Figure 3B{]}. These
community differences within a sample disappear when comparing 25 to 30
cycles and do not persist past 25 cycles {[}Figure 3{]}. Although these
trends are clearly noticeable, we found that there was no detectable
difference in Bray-Curtis index when comparing to the previous 5-cycle
increment for both fecal and mock communities after multiple comparison
correction (P-value \textgreater{} 0.05). Using PERMANOVA to test for
community differences based on cycle number within HiFi DNA polymerases,
only Phusion had cycle dependent differences (P-value = 0.03. For fecal
samples, Phusion was one of two HiFi DNA polymerases that had enough
sequences to be rarefied to 1000 at 15 cycles. Overall, these data
suggest that there are small HiFi DNA polymerase differences in
Bray-Curtis index that are dependent on cycle number.

\textbf{\emph{Sequence Error is Dependent on both HiFi DNA Polymerase
and Cycle Number:}} Differences in the median average per base error
varied by HiFi DNA polymerase without a clear pattern across
sub-sampling depth {[}Table S5{]}. The highest median average per base
error rates were for the Kappa HiFi DNA polymerase {[}Figure 4{]}. This
error rate was minimally affected by both the `pre.cluster' step and
chimera removal by VSEARCH {[}Figure 4{]}. The differences in the median
average per base error rate between the different HiFi DNA polymerases
was cycle dependent with Platinum having the largest changes versus
other HiFi DNA polymerases {[}Figure 4B-C and Table S6{]}. The total
sequences with at least one error was also cycle number dependent and
differences between HiFi DNA polymerases could be drastically reduced by
the use of the `pre.cluster' step {[}Figure S1{]}. These differences in
sequences with at least one error were mostly due to differences in
Accuprime\textsuperscript{TM} and Platinum versus the other HiFi DNA
polymerases {[}Figure S1 \& Table S7 \& S8{]}. Finally, we did not
observe a HiFi DNA polymerase dependent difference on base substitution
rate {[}Figure S2{]}. Although sequence error is dependent on HiFi DNA
polymerase some of these error dependent differences can be corrected
using existing bioinformatic approaches.

>>>>

\textbf{\emph{The Number of OTUs are Dependent on HiFi DNA Polymerase:}}
A consistent difference in the range of number of OTUs, that was
dependent on the HiFi DNA polymerase used was observed for fecal samples
{[}Figure 1{]}. Additionally, there was a trend for lower number of
cycles (15-20) to result in a reduced range in the number of OTUs versus
higher number of cycles (25, 30, and 35) between HiFi DNA polymerases
{[}Figure 1{]}. For fecal samples, there was a significant difference
between HiFi DNA polymerases at 35 cycles (F-stat \textgreater{} 16.35,
P-value = 9.7e-05) {[}Table S1{]}. No differences between the specific
groups was found based on a Tukey post-hoc test (P-value \textgreater{}
0.05) {[}Table S2{]}. This HiFi DNA polymerase dependent difference in
the range of number of OTUs also was observed in the mock community
samples {[}Figure 2{]}. Regardless if fecal or mock communities were
used, the same HiFi DNA polymerases had high (Platinum) and low
(Accuprime) number of OTUs and this was consistent across the number of
cycles used {[}Figure 1-2 \& Table S1-S4{]}. In contrast to the results
obtained with fecal samples, differences between HiFi DNA polymerases
were observed as early as 20 cycles in the mock community samples
(F-stat = 15.82, P-value = 0.002) {[}Table S3{]}. For the mock
community, the majority of differences in the number of OTUs were
between Kappa and Platinum and the other HiFi DNA polymerases across
different number of cycles {[}Table S4{]}. Based on these observations
in fecal and mock communities, it is clear that different HiFi DNA
polymerases result in a different total number of OTUs within a sample.

\textbf{\emph{Minimal Bray-Curtis Differences are Detected and are
Dependent on Cycle Number:}} A few small differences based on the number
of cycles were detected in overall bacterial community composition.
Within the same fecal sample and independent of HiFi DNA polymerases,
there were large differences in the community composition between 20 and
25 cycles {[}Figure 3A{]}. Additionally, when data was available for the
mock communities, there were also large observed differences between 20
and 25 cycles {[}Figure 3B{]}. These community differences within a
sample disappear when comparing 25 to 30 cycles and do not persist past
25 cycles {[}Figure 3{]}. Although these trends are clearly noticeable,
we found that there was no detectable difference in Bray-Curtis index
when comparing to the previous 5-cycle increment for both fecal and mock
communities after multiple comparison correction (P-value \textgreater{}
0.05). Using PERMANOVA to test for community differences based on the
number of cycles within HiFi DNA polymerases, only Phusion had cycle
dependent differences (P-value = 0.03. For fecal samples, Phusion was
one of two HiFi DNA polymerases that had enough sequences to be rarefied
to 1000 at 15 cycles. Overall, these data suggest that there are small
HiFi DNA polymerase differences in Bray-Curtis index that are dependent
on the number of cycles.

\textbf{\emph{Sequence Error is Dependent on both HiFi DNA Polymerase
and Cycle Number:}} Differences in the median error rate varied by HiFi
DNA polymerase across the number of cycles used {[}Table S5{]}. The
highest error rates were for the Kappa HiFi DNA polymerase {[}Figure
4{]}. The differences in the median error rate between the different
HiFi DNA polymerases was dependent on the number of cycles with Platinum
and Kappa having the most differences versus other HiFi DNA polymerases
{[}Figure 4 and Table S6{]}. The total sequences with at least one error
was also cycle number dependent {[}Table S7 \& S8. These differences in
sequences with at least one error were mostly due to differences in
Accuprime\textsuperscript{TM} and Kappa versus the other HiFi DNA
polymerases {[}Table S7 \& S8{]}. Finally, we did not observe a HiFi DNA
polymerase dependent difference on base substitution rate {[}Figure
S1{]}. These results suggest that sequence error is dependent on HiFi
DNA polymerase. We next wanted to assess whether chimeras also were
dependent on HiFi DNA polymerase and whether this affected OTU number.

### 2018-05-15 submission/supplemental.Rmd
remove sub-sampling and fix cycle jargon

mock_tukey <- as.data.frame(read_csv("../data/process/tables/mock_overall_tukey_results.csv")) %>% 
  mutate_at(vars(diff:upr), funs(round(., digits = 2))) %>% 
  filter(sub_sample_level == 1000) %>% 
  select(-sub_sample_level) %>% 
  rename(padj = `p adj`) %>% 
  mutate(padj = as.numeric(padj)) %>% 
  filter(padj < 0.05) %>% 
  arrange(padj) %>% 
  mutate_at(vars(padj), funs(format(., scientific = TRUE, digits = 2, nsmall = 2)))

@@

mock_error_rate_kruskal <- as.data.frame(read_csv("../data/process/tables/mock_error_overall_kruskal_results.csv")) %>% 
  mutate_at(vars(statistic), funs(round(., digits = 2))) %>% 
  mutate_at(vars(pvalue:bh), funs(format(., scientific = TRUE, digits = 2, nsmall = 2))) %>% 
  filter(sub_sample_level > 500)

colnames(mock_error_rate_kruskal) <- c("Rank Sum Statistic", "DF", "P-value", "BH", "Cycle", "Sub-Sample Depth")

@@

mock_error_rate_dunn <- as.data.frame(read_csv("../data/process/tables/mock_error_overall_dunn_results.csv")) %>% 
  mutate_at(vars(chi2:Z), funs(round(., digits = 2))) %>% 
  filter(sub_sample_level > 500, pvalue < 0.05) %>% 
  arrange(pvalue) %>% 
  mutate_at(vars(pvalue:bh), funs(format(., scientific = TRUE, digits = 2, nsmall = 2))) %>% 
  select(-bh)

colnames(mock_error_rate_dunn) <- c("Chi Squared", "Z Statistic", "P Adjusted", "Comparison", "Cycle", "Sub-Sample Depth")

@@

mock_error_count_kruskal <- as.data.frame(
  read_csv("../data/process/tables/mock_error_count_overall_kruskal_results.csv")) %>% 
  mutate_at(vars(statistic), funs(round(., digits = 2))) %>% 
  mutate_at(vars(pvalue:bh), funs(format(., scientific = TRUE, digits = 2, nsmall = 2))) %>% 
  filter(sub_sample_level > 500)

colnames(mock_error_count_kruskal) <- c("Rank Sum Statistic", "DF", "P-value", "BH", "Cycle", "Sub-Sample Depth")

@@

mock_error_count_dunn <- as.data.frame(read_csv("../data/process/tables/mock_error_count_overall_dunn_results.csv")) %>% 
  mutate_at(vars(chi2:Z), funs(round(., digits = 2))) %>% 
  filter(sub_sample_level > 500, pvalue < 0.05) %>% 
  arrange(pvalue) %>% 
  mutate_at(vars(pvalue:bh), funs(format(., scientific = TRUE, digits = 2, nsmall = 2))) %>% 
  select(-bh)

colnames(mock_error_count_dunn) <- c("Chi Squared", "Z Statistic", "P Adjusted", "Comparison", "Cycle", "Sub-Sample Depth")

>>>>

mock_tukey <- as.data.frame(read_csv("../data/process/tables/mock_overall_tukey_results.csv")) %>% 
  mutate_at(vars(diff:upr), funs(round(., digits = 2))) %>% 
  filter(sub_sample_level == 1000) %>% 
  select(-sub_sample_level) %>% 
  rename(padj = `p adj`) %>% 
  mutate(padj = as.numeric(padj)) %>% 
  filter(padj < 0.05) %>% 
  arrange(padj) %>% 
  mutate_at(vars(padj), funs(format(., scientific = TRUE, digits = 2, nsmall = 2))) %>% 
  arrange(cycle)

@@

mock_error_rate_kruskal <- as.data.frame(read_csv("../data/process/tables/mock_error_overall_kruskal_results.csv")) %>% 
  mutate_at(vars(statistic), funs(round(., digits = 2))) %>% 
  mutate_at(vars(pvalue:bh), funs(format(., scientific = TRUE, digits = 2, nsmall = 2))) %>% 
  select(-sub_sample_level)

colnames(mock_error_rate_kruskal) <- c("Rank Sum Statistic", "DF", "P-value", "BH", "Cycle")

@@

mock_error_rate_dunn <- as.data.frame(read_csv("../data/process/tables/mock_error_overall_dunn_results.csv")) %>% 
  mutate_at(vars(chi2:Z), funs(round(., digits = 2))) %>% 
  filter(pvalue < 0.05) %>% 
  arrange(pvalue) %>% 
  mutate_at(vars(pvalue:bh), funs(format(., scientific = TRUE, digits = 2, nsmall = 2))) %>% 
  select(-bh, -sub_sample_level) %>% 
  arrange(cycle)

colnames(mock_error_rate_dunn) <- c("Chi Squared", "Z Statistic", "P Adjusted", "Comparison", "Cycle")

@@

mock_error_count_kruskal <- as.data.frame(
  read_csv("../data/process/tables/mock_error_count_overall_kruskal_results.csv")) %>% 
  mutate_at(vars(statistic), funs(round(., digits = 2))) %>% 
  mutate_at(vars(pvalue:bh), funs(format(., scientific = TRUE, digits = 2, nsmall = 2))) %>% 
  select(-sub_sample_level)

colnames(mock_error_count_kruskal) <- c("Rank Sum Statistic", "DF", "P-value", "BH", "Cycle")

@@

mock_error_count_dunn <- as.data.frame(read_csv("../data/process/tables/mock_error_count_overall_dunn_results.csv")) %>% 
  mutate_at(vars(chi2:Z), funs(round(., digits = 2))) %>% 
  filter(pvalue < 0.05) %>% 
  arrange(pvalue) %>% 
  mutate_at(vars(pvalue:bh), funs(format(., scientific = TRUE, digits = 2, nsmall = 2))) %>% 
  select(-bh, -sub_sample_level) %>% 
  arrange(cycle)

colnames(mock_error_count_dunn) <- c("Chi Squared", "Z Statistic", "P Adjusted", "Comparison", "Cycle")
